Gaia's constraints on the extremely massive ($\sim$$150\,M_\odot$) runaway star candidate VFTS682

### 2018-06-10 VFTS682_APJL.tex
fixed bibliography

VFTS682 is identified with the source id 4657685637907503744 in the \emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band \citep{gaia:16,brown:18}.   The number of visibility periods, i.e. groups of observations separated from other groups by a gap of at least 4 days, used in the astrometric solution is 17 for this star. The reported astrometric excess noise is zero. \citep[][]{lindengren:18}.  These values suggest that the \emph{Gaia} DR2  data for VFTS682 are reliable. Generally the Gaia errors are correlated, but the correlation coefficient is small for this star, see Tab.~\ref{tab:vfts682}. We can therefore treat the proper motion components as uncorrelated.

To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.  They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both co-ordinates (see their  Sect.~2.1 for further discussion).

With these we compute the relative proper motion $\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.  We also derive the angle $\theta$ between the direction of motion and the vector connecting the center of R136 with the current position of VFTS 682,  i.e. such that $\theta = 0$ is the angle expected for a star ejected from R136.    We further provide the 2D velocity projected onto the radial vector pointing away from R136 $v_\mathrm{2d, \parallel}$  and the total projected 2d velocity $v_\mathrm{2d}$.   All kinematic quantities are provided in Tab.~\ref{tab:vfts682}.

@@

The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{Sabbi+2013}). Visually, the star is clearly isolated and does not appear to have any close neighbor. This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.

>>>>

VFTS682 is identified with the source id 4657685637907503744 in the
\emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band
\citep{gaia:16,brown:18}.   The number of visibility periods,
i.e. groups of observations separated from other groups by a gap of at
least 4 days, used in the astrometric solution is 17 for this
star. The reported astrometric excess noise is zero.  These values
suggest that the \emph{Gaia} DR2  data for VFTS682 are
reliable. Generally the Gaia errors are correlated  \citep[][]{lindengren:18}, but the correlation coefficients are small for this star% , see Tab.~\ref{tab:vfts682}
. We can therefore treat the proper motion components as uncorrelated.

To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.  They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).

With these we compute the relative proper motion
$\delta\mu_\mathrm{RA}$ and $\delta\mu_\mathrm{DEC}$.  We also derive
the angle $\theta$ between the direction of motion and the vector
connecting the center of R136 with the current position of VFTS 682,
i.e. such that $\theta = 0$ is the angle corresponding to perfectly
radial motion away from R136.    We further provide the 2D velocity
projected onto the radial vector pointing away from R136
$v_\mathrm{2d, \parallel}$  and the total projected 2d velocity
$v_\mathrm{2d}$.   All kinematic quantities are provided in
Tab.~\ref{tab:vfts682}. %%what does the projection on the proper
                        %%motion in the radial direction tell us?
                        %%Seems like a useless number to me...

@@

The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}). Visually, the star is clearly isolated and does not appear to have any close neighbor. This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.
